String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-05-05 2023-1-19/sections/algorithms.tex
solving CEFA

At first, we compute CEFA $\cB_{i} = \aut_{i,1} \cap \cdots \cap \aut_{i, l_i}$ for each $i \in [n]$. For each $i \in [n]$, let $\Lambda'_{x_i} = \{\cB_i\}$. 
Then the nonemptiness of CEFA in $\Lambda_{x_1}, \cdots, \Lambda_{x_n}$ w.r.t. $\varphi$ is reduced to the nonemptiness of CEFA in $\Lambda'_{x_1}, \cdots, \Lambda'_{x_n}$ w.r.t. $\varphi$.

@@

It turns out that to solve the nonemptiness of CEFA in $\Lambda'_{x_1}, \cdots, \Lambda'_{x_n}$ w.r.t. $\varphi$, we can ignore the characters in $\cB_1,\cdots, \cB_n$.

compute the product, ignore the alphabet, determinization, minimization

@@

>>>>

At first, we compute CEFA $\cB_{i} = \aut_{i,1} \cap \cdots \cap \aut_{i, l_i}$ for each $i \in [n]$. For each $i \in [n]$, let $\Lambda'_{x_i} = \{\cB_i\}$. Evidently, $R_{\cB_i}= \bigcup \limits_{j \in [l_i]} R_{\aut_{i, j}}$ for each $i \in [n]$.
Then the nonemptiness of CEFA in $\Lambda_{x_1}, \cdots, \Lambda_{x_n}$ w.r.t. $\varphi$ is reduced to the nonemptiness of CEFA in $\Lambda'_{x_1}, \cdots, \Lambda'_{x_n}$ w.r.t. $\varphi$.

@@

It turns out that to solve the nonemptiness of CEFA in $\Lambda'_{x_1}, \cdots, \Lambda'_{x_n}$ w.r.t. $\varphi$, we can ignore the characters in $\cB_1,\cdots, \cB_n$. This is due to the fact that there is only one CEFA for each string variable and $\varphi$ only cares about the values of registers.

%When the characters in $\cB_1, \cdots, \cB_n$ are ignored, they can be seen as NFAs. 
For each $i \in [n]$, let $\cC_i$ be the automaton obtained from $\cB_i$ by replacing each transition $(q, a, q', \myvec{v})$ with $(q, \myvec{v}, q')$. 
Then for each $i \in [n]$, $\cC_i$ can be seen as an NFA over the alphabet $V_i$, where $V_i$ is the set of vectors $\myvec{v}$ occurring in the transitions $(q, a, q', \myvec{v})$ of $\cB_i$.

Then the set of costs output by $\cB_i$, that is, $\{\myvec{v} \mid (w; \myvec{v}) \in \Lang(\cB_i)\}$ can be reduced to the computation of the Parikh image of $\cC_i$, namely, 
%
$$\{ \sum\limits_{\myvec{v} \in V_i} m_{\myvec{v}} \myvec{v} \mid \myvec{m} \in parikh(\cC_i)\}.$$

Moreover, the transitions $(q, \myvec{0}, q')$ in $\cC_i$ can be seen as $\varepsilon$-transitions, that is, we can replace each transition $(q, \myvec{0}, q')$ with $(q, \varepsilon, q')$, without affecting the set $\{ \sum\limits_{\myvec{v} \in V_i} m_{\myvec{v}} \myvec{v} \mid \myvec{m} \in parikh(\cC_i)\}$. Therefore, we first do these replacements, apply the $\varepsilon$-transition removal procedure. Then we utilize the determinization and minimization algorithm for NFA to reduce the size of $\cC_i$.

Finally, if there are two indices $1 \le i' < j' \le |R_{\cB_i}|$ such that $v_{i'} = v_{j'}$ for each vector $\myvec{v} \in V_i$, then we can remove the $j'$-th component from the characters (vectors) of $\cC_i$, and let $\varphi := \varphi \wedge r_{i, i'} = r_{i, j'}$. By abusing the notation, we still use $\cC_i$ to denote the resulting NFA.

By doing this, we reduce the automata sizes. Then we compute an LIA formula $\psi_{i}$ from $\cC_i$ to specify the Parikh images of $\cC_i$.  Then the nonemptiness of $\cB_1, \cdots, \cB_n$ w.r.t. $\varphi$ is reduced to the satisfiability of $\psi_1 \wedge \cdots \psi_n \wedge \varphi$.

@@

In summary, we obtain an algorithm for solving the nonemptiness of $\Lambda_{x_1}, \cdots, \Lambda_{x_n}$ w.r.t. $\varphi$, as illustrated in Figure~\ref{xxx}.
